### 2019-09-11 reports/thesis/chapters/3_compoundPoisson.tex
Use proofreader

In an x-ray tube, photons are emitted as a Poisson process \citep{whiting2006properties, cierniak2011x} and each photon has some random energy due to bremsstrahlung and characteristic radiation \citep{sun2012overview}. This shares similarities to the compound Poisson distribution. Let $Y$ be the number of photons emitted for some time exposure $\tau$, then $Y\sim\poisson(\lambda)$. Each photon is assumed to be i.i.d.~with random energy $U_i$ for $i=1,2,3,\dotsc$ with p.d.f.~$p_U(u)$. The random variables discussed here cover all the latent variables in the compound Poisson.

@@

Experiments have been done to verify the compound Poisson nature of the detector. This was done by investigating the variance of radiographs of air \citep{hsieh2015compound} and a polyethylene cylinder \citep{yang2009evaluation, yang2010noise} using different exposures. It was found there were two components in the noise, one was signal dependent and comes from the compound Poisson distribution, the other was signal independent and may be electronic noise. The electronic noise can be modelled using a Normal distribution \citep{xu2009electronic}.

@@

One special property of the compound Poisson-gamma distribution is that it is in the Tweedie dispersion exponential family \citep{jorgensen1987exponential}. It can be shown that it has a special variance mean relationship
\begin{equation}
  \variance[X] = \phi \mu^p
\end{equation}
where $1<p<2$. This is derived in Appendix \ref{chapter:appendix_tweedie}. It should be noted that this relationship is for fixed $p$ and $\phi$. This is different from the linear variance mean relationship found at the start of the chapter which was for fixed $\alpha$ and $\beta$ from assuming no beam hardening.

@@

The evaluation of the p.d.f.~of a compound Poisson-gamma distribution is useful so that the likelihood can be obtained. The likelihood can be used to find, for example, maximum likelihood estimators. A problem occurs when dealing with the infinite sum in the p.d.f.~because it cannot be simplified. There are a number of approximations or computational methods to evaluate the p.d.f.~such as Fourier inverting the characteristic function \citep{dunn2008evaluation}, using the saddlepoint approximation \citep{daniels1954saddlepoint} or cleverly sum over certain terms in the infinite sum \citep{dunn2005series}. Monte Carlo methods can be used to evaluate the p.d.f.~by simulating compound Poisson-gamma random variables.

@@

The infinite sum, $\sum_{y=1}^\infty W_y$, can be computationally summed in a clever way in order to evaluate the p.d.f. This was done by summing only large terms in the sum and ignoring small terms \citep{dunn2005series}. \cite{dunn2005series} approximated the sum by truncation
\begin{equation}
  \sum_{y=1}^\infty W_y \approx \sum_{y=y_\text{l}}^{y_\text{u}}W_y
\end{equation}
where $y_\text{l}<y_{\text{max}}<y_\text{u}$ and $y_{\text{max}}$ is the value of $y$ which maximises $W_y$. \cite{dunn2005series} used Stirling's approximation to find that
\begin{equation}
  y_{\text{max}} \approx \frac{x^{2-p}}{\phi(2-p)}
\end{equation}
by treating $W_y$ as a continuous and differentiable function of $y$. The derivation is shown in Appendix \ref{chapter:appendix_compoundPoissonSeries}. Because values of $y$ are positive integers, it would be appropriate to round $y_\text{max}$ accordingly
\begin{equation}
  y_{\text{max}} = \text{max}\left[
    1,\text{round}\left(\frac{x^{2-p}}{\phi(2-p)}\right)
  \right]
  \ .
\end{equation}

@@

Simulations of a compound Poisson-gamma random variable were conducted with the aim to compare how well these density evaluation methods performed. This was done by comparing the evaluated densities using the histogram of simulations and a Q-Q plot. The following compound Poisson-gamma distributions were used in the simulations to capture the variety in the compound Poisson-gamma family: $\CPoisson(1,1,1)$, $\CPoisson(1,100,1)$, $\CPoisson(10,1,1)$, $\CPoisson(100,100,1)$. For each of these distribution, 1\,000 random variables were simulated. Varying $\beta$ is not interesting as this only scales the random variable.

There were a few technical problems with the histogram because the compound Poisson has probability mass at zero and probability density for positive numbers. To correctly represent the empirical density of a compound Poisson random variable, a bar chart was used to show the frequency of a zero and a histogram to show the frequency density of positive numbers. However, the Normal approximation and the saddlepoint approximation does not have mass support at zero. Therefore, to compare these approximate densities to the empirical distribution fairly, a histogram containing both zero and positive samples was used when appropriate.

@@

Let $\widehat{\lambda}$, $\widehat{\alpha}$ and $\widehat{\beta}$ be estimators of $\lambda$, $\alpha$ and $\beta$ respectively. The log likelihood is defined to be
\begin{equation*}
  \ln L(\lambda,\alpha,\beta;X) = \sum_{i=1}^n
  \left[
    \mathbb{I}(x_i=0)\ln \prob(X=x_i)
    +\mathbb{I}(x_i>0)\ln p_X(x_i)
  \right]
\end{equation*}
and using the results in Appendix \ref{chapter:appendix_tweedie}
\begin{equation}
  \ln L(\lambda,\alpha,\beta;X) =
    \sum_{i=1}^n
    \left[
      -\mathbb{I}(x_i=0)\lambda
      \vphantom{\sum_i^i}
    \right.
    {}\\
    \left.
      +\mathbb{I}(x_i>0)\left(
      -\beta x_i-\lambda-\ln x +\ln\sum_{y=1}^\infty W_y
      \right)
    \right] \ .
\end{equation}
Maximum likelihood estimators are values of $\lambda$, $\alpha$ and $\beta$ which jointly maximise the log likelihood.

The EM algorithm \citep{dempster1977maximum} treats $Y\sim\poisson(\lambda)$ as a latent variable. Let $\left\{Y_1,Y_2,Y_3,\dotsc, Y_n\right\}$ be realisations of $Y$. Define the joint log likelihood to be
\begin{multline*}
  \ln L(\lambda,\alpha,\beta;X,Y)
  =
  \sum_{i=1}^n
  \left[
    \mathbb{I}(x_i=0)
    \ln
    \prob(Y=0)
  \right.
  \\
  \left.
    +
    \mathbb{I}(x_i>0)
    \ln
    \left[
      p_{X|Y}(x_i|Y_i)\prob(Y=Y_i)
    \right]
  \right]
\end{multline*}
so that
\begin{multline}
  \ln L(\lambda,\alpha,\beta;X,Y)=
  \sum_{i=1}^n
  \left[
    -\mathbb{I}(x_i=0)
    \lambda
  \right.
  \\
  \left.+
    \mathbb{I}(x_i>0)
    \left(
      Y_i\alpha\ln\beta-\ln\Gamma(Y_i\alpha)+(Y_i\alpha-1)\ln x_i - \beta x_i
    \right.
  \right.
  \\
  \left.
    \left.  
      - \lambda + Y_i \ln \lambda - \ln(Y_i!)
    \right)
  \right]
  \ .
\end{multline}
The estimators were found by optimising the joint log likelihood. This was done iteratively by estimating the $Y$s given the $X$s and parameters (E step), followed by estimating the parameters given the $Y$s and $X$s (M step) until some convergence conditions were met.

@@

In the M step, the conditional expected joint log likelihood is maximised with respect to the parameters $\lambda$, $\alpha$ and $\beta$. The objective function is
\begin{align}
  T(\lambda,\alpha,\beta)&=
  \begin{multlined}[t]
    \sum_{i=1}^n
    \expectation\left[
      -\mathbb{I}(x_i=0)
      \lambda
    \right.
    \\
    \left.+
      \mathbb{I}(x_i>0)
      \left(
        Y_i\alpha\ln\beta-\ln\Gamma(Y_i\alpha)+(Y_i\alpha-1)\ln x_i - \beta x_i
      \right.
    \right.
    \\
    \left.
      \left.  
        - \lambda + Y_i \ln \lambda - \ln(Y_i!)
      \right)
      |X_i=x_i
    \right]
  \end{multlined}
  \nonumber\\
  &\!\begin{multlined}[b]=
    -n\lambda
    \\
    +\sum_{i=1}^n
    \mathbb{I}(x_i>0)
    \left[
      \expectation[Y_i|X_i=x_i]\alpha\ln\beta-\expectation[\ln\Gamma(Y_i\alpha)|X_i=x_i]
    \right.
    {}\\
    \left.
      +\expectation[Y_i|X_i=x_i]\alpha\ln x_i - \beta x_i
      + \expectation[Y_i|X_i=x_i] \ln \lambda
    \right] + c
  \end{multlined}
\end{align}
where $c$ is some constant not dependent on $\lambda$, $\alpha$ or $\beta$.

@@

Maximising $T$ with respect to $\alpha$ and $\beta$ can be done numerically using the Newton-Raphson method since derivatives up to the second order can be obtained. For the first order, these are:
\begin{multline}
  \frac{\partial T}{\partial \alpha} =
  \sum_{i=1}^n\mathbb{I}(x_i>0)
  \left[
    y_i\ln\beta -\psi(\alpha y_i)y_i-\zeta_i\alpha\psi'(\alpha y_i)
    \vphantom{\frac{1}{1}}
  \right.
  \\
  \left.
    - \frac{1}{2}\zeta_i\alpha^2\psi''(\alpha y_i)y_i + y_i\ln x_i
  \right]
\end{multline}
and
\begin{equation}
  \frac{\partial T}{\partial \beta} = \sum_{i=1}^n\mathbb{I}(x_i>0)\left[
  \frac{\alpha y_i}{\beta}-x_i
  \right]
  \ .
\end{equation}
The second orders are:
\begin{equation}
  \frac{\partial^2 T}{\partial \alpha \partial \beta} =
  \sum_{i=1}^n \mathbb{I}(x_i>0)\left[\frac{y_i}{\beta}\right]
  \ ,
  \label{eq:compoundPoisson:d2tdadb}
\end{equation}
\begin{equation}
  \frac{\partial^2 T}{\partial \beta^2} = \sum_{i=1}^n\mathbb{I}(x_i>0)\left[-\frac{\alpha y_i}{\beta^2}
  \right]
  \label{eq:compoundPoisson:d2td2b}
\end{equation}
and
\begin{multline*}
  \frac{\partial^2 T}{\partial \alpha^2} = 
  \sum_{i=1}^n\mathbb{I}(x_i>0)
  \left[
    -y_i^2\psi'(\alpha y_i) - \zeta_i\psi'(\alpha y_i) - \zeta_i\alpha y_i\psi''(\alpha y_i)
    \vphantom{\frac{1}{2}}\right.\\\left. 
    - \zeta_i\alpha\psi''(\alpha y_i)y_i
    -\frac{1}{2}\zeta_i\alpha^2\psi'''(\alpha y_i)y_i^2
  \right]
\end{multline*}
simplifying to
\begin{multline}
  \frac{\partial^2 T}{\partial \alpha^2} = 
  \sum_{i=1}^n\mathbb{I}(x_i>0)
  \left[
    -(y_i^2+\zeta_i)\psi'(\alpha y_i) - 2\zeta_i\alpha y_i\psi''(\alpha y_i)
    \vphantom{\frac{1}{2}}
  \right.
  \\
  \left.  
    -\frac{1}{2}\zeta_i\alpha^2y_i^2\psi'''(\alpha y_i)
  \right] \ .
  \label{eq:compoundPoisson:d2td2a}
\end{multline}

@@

An experiment was conducted to assess the performance on the EM algorithm. For a given set of parameters, 1\,000 samples of a $\CPoisson(\lambda,\alpha,\beta)$ random variable were simulated. The EM algorithm was initialised with its parameters at the true values to investigate the convergence in that vicinity. The log likelihood $\ln L(\lambda,\alpha,\beta;X)$ and the parameters were recorded at every EM step. The experiment was repeated 10 times using different simulated samples.

@@

For $\lambda=10$ and $\lambda=100$ case, the estimates of $\alpha$ and $\beta$ struggled to converge and increased/decreased without bounds without affecting the log likelihood. It appeared that the EM algorithm failed for $\lambda>10$ looking at these particular examples.

\begin{figure}[p]
  \centering
    \centerline{
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/CpEmAlgorithm_CompoundPoissonlambda1alpha1beta1_lnL.eps}
        \caption{Log likelihood}
    \end{subfigure}
        \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/CpEmAlgorithm_CompoundPoissonlambda1alpha1beta1_lambda.eps}
        \caption{$\lambda$}
    \end{subfigure}
    }
    \centerline{
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/CpEmAlgorithm_CompoundPoissonlambda1alpha1beta1_alpha.eps}
        \caption{$\alpha$}
    \end{subfigure}
        \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/CpEmAlgorithm_CompoundPoissonlambda1alpha1beta1_beta.eps}
        \caption{$\beta$}
    \end{subfigure}
    }
    \caption{EM algorithm was used to estimate the parameters of a $\CPoisson(1,1,1)$ random variable using 1\,000 simulated samples, repeated 10 times. The graphs show the log likelihood and the estimated parameters at each EM step for each repeat of the experiment. The dashed lines show the standard deviation using the Cram\'er-Rao lower bound.}
    \label{fig:compoundPoisson_convergence_1}
\end{figure}

\begin{figure}[p]
  \centering
    \centerline{
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/CpEmAlgorithm_CompoundPoissonlambda1alpha100beta1_lnL.eps}
        \caption{Log likelihood}
    \end{subfigure}
        \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/CpEmAlgorithm_CompoundPoissonlambda1alpha100beta1_lambda.eps}
        \caption{$\lambda$}
    \end{subfigure}
    }
    \centerline{
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/CpEmAlgorithm_CompoundPoissonlambda1alpha100beta1_alpha.eps}
        \caption{$\alpha$}
    \end{subfigure}
        \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/CpEmAlgorithm_CompoundPoissonlambda1alpha100beta1_beta.eps}
        \caption{$\beta$}
    \end{subfigure}
    }
    \caption{EM algorithm was used to estimate the parameters of a $\CPoisson(1,100,1)$ random variable using 1\,000 simulated samples, repeated 10 times. The graphs show the log likelihood and the estimated parameters at each EM step for each repeat of the experiment. The dashed lines show the standard deviation using the Cram\'er-Rao lower bound.}
    \label{fig:compoundPoisson_convergence_2}
\end{figure}

\begin{figure}[p]
  \centering
    \centerline{
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/CpEmAlgorithm_CompoundPoissonlambda10alpha1beta1_lnL.eps}
        \caption{Log likelihood}
    \end{subfigure}
        \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/CpEmAlgorithm_CompoundPoissonlambda10alpha1beta1_lambda.eps}
        \caption{$\lambda$}
    \end{subfigure}
    }
    \centerline{
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/CpEmAlgorithm_CompoundPoissonlambda10alpha1beta1_alpha.eps}
        \caption{$\alpha$}
    \end{subfigure}
        \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/CpEmAlgorithm_CompoundPoissonlambda10alpha1beta1_beta.eps}
        \caption{$\beta$}
    \end{subfigure}
    }
    \caption{EM algorithm was used to estimate the parameters of a $\CPoisson(10,1,1)$ random variable using 1\,000 simulated samples, repeated 10 times. The graphs show the log likelihood and the estimated parameters at each EM step for each repeat of the experiment. The dashed lines show the standard deviation using the Cram\'er-Rao lower bound.}
    \label{fig:compoundPoisson_convergence_3}
\end{figure}

\begin{figure}[p]
  \centering
    \centerline{
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/CpEmAlgorithm_CompoundPoissonlambda100alpha100beta1_lnL.eps}
        \caption{Log likelihood}
    \end{subfigure}
        \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/CpEmAlgorithm_CompoundPoissonlambda100alpha100beta1_lambda.eps}
        \caption{$\lambda$}
    \end{subfigure}
    }
    \centerline{
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/CpEmAlgorithm_CompoundPoissonlambda100alpha100beta1_alpha.eps}
        \caption{$\alpha$}
    \end{subfigure}
        \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/CpEmAlgorithm_CompoundPoissonlambda100alpha100beta1_beta.eps}
        \caption{$\beta$}
    \end{subfigure}
    }
    \caption{EM algorithm was used to estimate the parameters of a $\CPoisson(100,100,1)$ random variable using 1\,000 simulated samples, repeated 10 times. The graphs show the log likelihood and the estimated parameters at each EM step for each repeat of the experiment. The dashed lines show the standard deviation using the Cram\'er-Rao lower bound.}
    \label{fig:compoundPoisson_convergence_4}
\end{figure}

@@

\subsection{Log Likelihood Plot}

>>>>

In an x-ray tube, photons are emitted as a Poisson process \citep{whiting2006properties, cierniak2011x} and each photon has some random energy due to bremsstrahlung and characteristic radiation \citep{sun2012overview}. This is similar to the compound Poisson distribution. Let $Y$ be the number of photons emitted for some time exposure $\tau$, then $Y\sim\poisson(\lambda)$. Each photon is assumed to be i.i.d.~with random energy $U_i$ for $i=1,2,3,\dotsc$ with p.d.f.~$p_U(u)$. The random variables discussed here cover all the latent variables in the compound Poisson.

@@

Experiments have been done to verify the compound Poisson nature of the detector. This was done by investigating the variance of radiographs of air \citep{hsieh2015compound} and a polyethene cylinder \citep{yang2009evaluation, yang2010noise} using different exposures. It was found there were two components in the noise, one was signal-dependent and comes from the compound Poisson distribution, the other was signal independent and maybe electronic noise. The electronic noise can be modelled using a Normal distribution \citep{xu2009electronic}.

@@

One special property of the compound Poisson-gamma distribution is that it is in the Tweedie dispersion exponential family \citep{jorgensen1987exponential}. It can be shown that it has a special variance-mean relationship
\begin{equation}
  \variance[X] = \phi \mu^p
\end{equation}
where $1<p<2$. This is derived in Appendix \ref{chapter:appendix_tweedie}. It should be noted that this relationship is for fixed $p$ and $\phi$. This is different from the linear variance-mean relationship found at the start of the chapter which was for fixed $\alpha$ and $\beta$ from assuming no beam hardening.

@@

The evaluation of the p.d.f.~of a compound Poisson-gamma distribution is useful so that the likelihood can be obtained. The likelihood can be used to find, for example, maximum likelihood estimators. A problem occurs when dealing with the infinite sum in the p.d.f.~because it cannot be simplified. There are several approximations or computational methods to evaluate the p.d.f.~such as Fourier inverting the characteristic function \citep{dunn2008evaluation}, using the saddlepoint approximation \citep{daniels1954saddlepoint} or cleverly sum over certain terms in the infinite sum \citep{dunn2005series}. Monte Carlo methods can be used to evaluate the p.d.f.~by simulating compound Poisson-gamma random variables.

@@

The infinite sum, $\sum_{y=1}^\infty W_y$, can be computationally summed cleverly to evaluate the p.d.f. This was done by summing only large terms in the sum and ignoring small terms \citep{dunn2005series}. \cite{dunn2005series} approximated the sum by truncation
\begin{equation}
  \sum_{y=1}^\infty W_y \approx \sum_{y=y_\text{l}}^{y_\text{u}}W_y
\end{equation}
where $y_\text{l}<y_{\text{max}}<y_\text{u}$ and $y_{\text{max}}$ is the value of $y$ which maximises $W_y$. \cite{dunn2005series} used Stirling's approximation to find that
\begin{equation}
  y_{\text{max}} \approx \frac{x^{2-p}}{\phi(2-p)}
\end{equation}
by treating $W_y$ as a continuous and differentiable function of $y$. The derivation is shown in Appendix \ref{chapter:appendix_compoundPoissonSeries}. Because values of $y$ are positive integers, it would be appropriate to round $y_\text{max}$ accordingly
\begin{equation}
  y_{\text{max}} = \text{max}\left[
    1,\text{round}\left(\frac{x^{2-p}}{\phi(2-p)}\right)
  \right]
  \ .
\end{equation}

@@

Simulations of a compound Poisson-gamma random variable were conducted to compare how well these density evaluation methods performed. This was done by comparing the evaluated densities using the histogram of simulations and a Q-Q plot. The following compound Poisson-gamma distributions were used in the simulations to capture the variety in the compound Poisson-gamma family: $\CPoisson(1,1,1)$, $\CPoisson(1,100,1)$, $\CPoisson(10,1,1)$, $\CPoisson(100,100,1)$. For each of these distribution, 1\,000 random variables were simulated. Varying $\beta$ is not interesting as this only scales the random variable.

There were a few technical problems with the histogram because the compound Poisson has probability mass at zero and probability density for positive numbers. To correctly represent the empirical density of a compound Poisson random variable, a bar chart was used to show the frequency of zeros and a histogram to show the frequency density of positive numbers. However, the Normal approximation and the saddlepoint approximation does not have mass support at zero. Therefore, to compare these approximate densities to the empirical distribution fairly, a histogram containing both zero and positive samples was used when appropriate.

@@

Let $\widehat{\lambda}$, $\widehat{\alpha}$ and $\widehat{\beta}$ be estimators of $\lambda$, $\alpha$ and $\beta$ respectively. The log-likelihood is defined to be
\begin{equation*}
  \ln L(\lambda,\alpha,\beta;X) = \sum_{i=1}^n
  \left[
    \mathbb{I}(x_i=0)\ln \prob(X=x_i)
    +\mathbb{I}(x_i>0)\ln p_X(x_i)
  \right]
\end{equation*}
and using the results in Appendix \ref{chapter:appendix_tweedie}
\begin{equation}
  \ln L(\lambda,\alpha,\beta;X) =
    \sum_{i=1}^n
    \left[
      -\mathbb{I}(x_i=0)\lambda
      \vphantom{\sum_i^i}
    \right.
    {}\\
    \left.
      +\mathbb{I}(x_i>0)\left(
      -\beta x_i-\lambda-\ln x +\ln\sum_{y=1}^\infty W_y
      \right)
    \right] \ .
\end{equation}
Maximum likelihood estimators are values of $\lambda$, $\alpha$ and $\beta$ which jointly maximise the log-likelihood.

The EM algorithm \citep{dempster1977maximum} treats $Y\sim\poisson(\lambda)$ as a latent variable. Let $\left\{Y_1,Y_2,Y_3,\dotsc, Y_n\right\}$ be realisations of $Y$. Define the joint log-likelihood to be
\begin{multline*}
  \ln L(\lambda,\alpha,\beta;X,Y)
  =
  \sum_{i=1}^n
  \left[
    \mathbb{I}(x_i=0)
    \ln
    \prob(Y=0)
  \right.
  \\
  \left.
    +
    \mathbb{I}(x_i>0)
    \ln
    \left[
      p_{X|Y}(x_i|Y_i)\prob(Y=Y_i)
    \right]
  \right]
\end{multline*}
so that
\begin{multline}
  \ln L(\lambda,\alpha,\beta;X,Y)=
  \sum_{i=1}^n
  \left[
    -\mathbb{I}(x_i=0)
    \lambda
  \right.
  \\
  \left.+
    \mathbb{I}(x_i>0)
    \left(
      Y_i\alpha\ln\beta-\ln\Gamma(Y_i\alpha)+(Y_i\alpha-1)\ln x_i - \beta x_i
    \right.
  \right.
  \\
  \left.
    \left.  
      - \lambda + Y_i \ln \lambda - \ln(Y_i!)
    \right)
  \right]
  \ .
\end{multline}
The estimators were found by optimising the joint log-likelihood. This was done iteratively by estimating the $Y$s given the $X$s and parameters (E step), followed by estimating the parameters given the $Y$s and $X$s (M step) until some convergence conditions were met.

@@

In the M step, the conditional expected joint log-likelihood is maximised with respect to the parameters $\lambda$, $\alpha$ and $\beta$. The objective function is
\begin{align}
  T(\lambda,\alpha,\beta)&=
  \begin{multlined}[t]
    \sum_{i=1}^n
    \expectation\left[
      -\mathbb{I}(x_i=0)
      \lambda
    \right.
    \\
    \left.+
      \mathbb{I}(x_i>0)
      \left(
        Y_i\alpha\ln\beta-\ln\Gamma(Y_i\alpha)+(Y_i\alpha-1)\ln x_i - \beta x_i
      \right.
    \right.
    \\
    \left.
      \left.  
        - \lambda + Y_i \ln \lambda - \ln(Y_i!)
      \right)
      |X_i=x_i
    \right]
  \end{multlined}
  \nonumber\\
  &\!\begin{multlined}[b]=
    -n\lambda
    \\
    +\sum_{i=1}^n
    \mathbb{I}(x_i>0)
    \left[
      \expectation[Y_i|X_i=x_i]\alpha\ln\beta-\expectation[\ln\Gamma(Y_i\alpha)|X_i=x_i]
    \right.
    {}\\
    \left.
      +\expectation[Y_i|X_i=x_i]\alpha\ln x_i - \beta x_i
      + \expectation[Y_i|X_i=x_i] \ln \lambda
    \right] + c
  \end{multlined}
\end{align}
where $c$ is some constant not dependent on $\lambda$, $\alpha$ or $\beta$.

@@

Maximising $T$ with respect to $\alpha$ and $\beta$ can be done numerically using the Newton-Raphson method since derivatives up to the second-order can be obtained. For the first-order derivatives, these are:
\begin{multline}
  \frac{\partial T}{\partial \alpha} =
  \sum_{i=1}^n\mathbb{I}(x_i>0)
  \left[
    y_i\ln\beta -\psi(\alpha y_i)y_i-\zeta_i\alpha\psi'(\alpha y_i)
    \vphantom{\frac{1}{1}}
  \right.
  \\
  \left.
    - \frac{1}{2}\zeta_i\alpha^2\psi''(\alpha y_i)y_i + y_i\ln x_i
  \right]
\end{multline}
and
\begin{equation}
  \frac{\partial T}{\partial \beta} = \sum_{i=1}^n\mathbb{I}(x_i>0)\left[
  \frac{\alpha y_i}{\beta}-x_i
  \right]
  \ .
\end{equation}
The second-order derivatives are:
\begin{equation}
  \frac{\partial^2 T}{\partial \alpha \partial \beta} =
  \sum_{i=1}^n \mathbb{I}(x_i>0)\left[\frac{y_i}{\beta}\right]
  \ ,
  \label{eq:compoundPoisson:d2tdadb}
\end{equation}
\begin{equation}
  \frac{\partial^2 T}{\partial \beta^2} = \sum_{i=1}^n\mathbb{I}(x_i>0)\left[-\frac{\alpha y_i}{\beta^2}
  \right]
  \label{eq:compoundPoisson:d2td2b}
\end{equation}
and
\begin{multline*}
  \frac{\partial^2 T}{\partial \alpha^2} = 
  \sum_{i=1}^n\mathbb{I}(x_i>0)
  \left[
    -y_i^2\psi'(\alpha y_i) - \zeta_i\psi'(\alpha y_i) - \zeta_i\alpha y_i\psi''(\alpha y_i)
    \vphantom{\frac{1}{2}}\right.\\\left. 
    - \zeta_i\alpha\psi''(\alpha y_i)y_i
    -\frac{1}{2}\zeta_i\alpha^2\psi'''(\alpha y_i)y_i^2
  \right]
\end{multline*}
simplifying to
\begin{multline}
  \frac{\partial^2 T}{\partial \alpha^2} = 
  \sum_{i=1}^n\mathbb{I}(x_i>0)
  \left[
    -(y_i^2+\zeta_i)\psi'(\alpha y_i) - 2\zeta_i\alpha y_i\psi''(\alpha y_i)
    \vphantom{\frac{1}{2}}
  \right.
  \\
  \left.  
    -\frac{1}{2}\zeta_i\alpha^2y_i^2\psi'''(\alpha y_i)
  \right] \ .
  \label{eq:compoundPoisson:d2td2a}
\end{multline}

@@

An experiment was conducted to assess the performance on the EM algorithm. For a given set of parameters, 1\,000 samples of a $\CPoisson(\lambda,\alpha,\beta)$ random variable were simulated. The EM algorithm was initialised with its parameters at the true values to investigate the convergence in that vicinity. The log-likelihood $\ln L(\lambda,\alpha,\beta;X)$ and the parameters were recorded at every EM step. The experiment was repeated 10 times using different simulated samples.

@@

For $\lambda=10$ and $\lambda=100$ case, the estimates of $\alpha$ and $\beta$ struggled to converge and increased/decreased without bounds without affecting the log-likelihood. It appeared that the EM algorithm failed for $\lambda>10$ looking at these particular examples.

\begin{figure}[p]
  \centering
    \centerline{
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/CpEmAlgorithm_CompoundPoissonlambda1alpha1beta1_lnL.eps}
        \caption{Log-likelihood}
    \end{subfigure}
        \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/CpEmAlgorithm_CompoundPoissonlambda1alpha1beta1_lambda.eps}
        \caption{$\lambda$}
    \end{subfigure}
    }
    \centerline{
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/CpEmAlgorithm_CompoundPoissonlambda1alpha1beta1_alpha.eps}
        \caption{$\alpha$}
    \end{subfigure}
        \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/CpEmAlgorithm_CompoundPoissonlambda1alpha1beta1_beta.eps}
        \caption{$\beta$}
    \end{subfigure}
    }
    \caption{EM algorithm was used to estimate the parameters of a $\CPoisson(1,1,1)$ random variable using 1\,000 simulated samples, repeated 10 times. The graphs show the log-likelihood and the estimated parameters at each EM step for each repeat of the experiment. The dashed lines show the standard deviation using the Cram\'er-Rao lower bound.}
    \label{fig:compoundPoisson_convergence_1}
\end{figure}

\begin{figure}[p]
  \centering
    \centerline{
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/CpEmAlgorithm_CompoundPoissonlambda1alpha100beta1_lnL.eps}
        \caption{Log-likelihood}
    \end{subfigure}
        \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/CpEmAlgorithm_CompoundPoissonlambda1alpha100beta1_lambda.eps}
        \caption{$\lambda$}
    \end{subfigure}
    }
    \centerline{
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/CpEmAlgorithm_CompoundPoissonlambda1alpha100beta1_alpha.eps}
        \caption{$\alpha$}
    \end{subfigure}
        \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/CpEmAlgorithm_CompoundPoissonlambda1alpha100beta1_beta.eps}
        \caption{$\beta$}
    \end{subfigure}
    }
    \caption{EM algorithm was used to estimate the parameters of a $\CPoisson(1,100,1)$ random variable using 1\,000 simulated samples, repeated 10 times. The graphs show the log-likelihood and the estimated parameters at each EM step for each repeat of the experiment. The dashed lines show the standard deviation using the Cram\'er-Rao lower bound.}
    \label{fig:compoundPoisson_convergence_2}
\end{figure}

\begin{figure}[p]
  \centering
    \centerline{
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/CpEmAlgorithm_CompoundPoissonlambda10alpha1beta1_lnL.eps}
        \caption{Log-likelihood}
    \end{subfigure}
        \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/CpEmAlgorithm_CompoundPoissonlambda10alpha1beta1_lambda.eps}
        \caption{$\lambda$}
    \end{subfigure}
    }
    \centerline{
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/CpEmAlgorithm_CompoundPoissonlambda10alpha1beta1_alpha.eps}
        \caption{$\alpha$}
    \end{subfigure}
        \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/CpEmAlgorithm_CompoundPoissonlambda10alpha1beta1_beta.eps}
        \caption{$\beta$}
    \end{subfigure}
    }
    \caption{EM algorithm was used to estimate the parameters of a $\CPoisson(10,1,1)$ random variable using 1\,000 simulated samples, repeated 10 times. The graphs show the log-likelihood and the estimated parameters at each EM step for each repeat of the experiment. The dashed lines show the standard deviation using the Cram\'er-Rao lower bound.}
    \label{fig:compoundPoisson_convergence_3}
\end{figure}

\begin{figure}[p]
  \centering
    \centerline{
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/CpEmAlgorithm_CompoundPoissonlambda100alpha100beta1_lnL.eps}
        \caption{Log-likelihood}
    \end{subfigure}
        \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/CpEmAlgorithm_CompoundPoissonlambda100alpha100beta1_lambda.eps}
        \caption{$\lambda$}
    \end{subfigure}
    }
    \centerline{
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/CpEmAlgorithm_CompoundPoissonlambda100alpha100beta1_alpha.eps}
        \caption{$\alpha$}
    \end{subfigure}
        \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/CpEmAlgorithm_CompoundPoissonlambda100alpha100beta1_beta.eps}
        \caption{$\beta$}
    \end{subfigure}
    }
    \caption{EM algorithm was used to estimate the parameters of a $\CPoisson(100,100,1)$ random variable using 1\,000 simulated samples, repeated 10 times. The graphs show the log-likelihood and the estimated parameters at each EM step for each repeat of the experiment. The dashed lines show the standard deviation using the Cram\'er-Rao lower bound.}
    \label{fig:compoundPoisson_convergence_4}
\end{figure}

@@

\subsection{Log-Likelihood Plot}
